---
title: "HW 1 Data Science"
author: "Victoria Mello (vsm2118)"
date: "September 19, 2023"
output: github_document
---
# Promblem 1

```{r}
# Load the moderndive package
library(moderndive)
library(tidyverse)

# Load the early_january_weather dataset
data("early_january_weather")
```

```{r}
# Access dataset help file
?early_january_weather
```
The dataset “early_january_weather” has 15 variables total:

* origin (weather station)

* year, month, day, hour (time of recording)

* temp, dewp (temperature and dewpoint in F)

* humid (relative humidity)

* wind_dir, wind_speed, wind_gust (wind direction (in degrees), speed and gust speed (in mph))

* precip (precipitation in inches)

* pressure (sea level pressure in millibars)

* visib (visibility in miles)

* ftime_hour (date and hour of the recording)

The most important variables in this dataset for this analysis are temp, humid, and time_hour. There are `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns in the dataset. The mean temperature in the dataset is `r mean(early_january_weather[["temp"]])` degrees fahrenheit.

```{r}
# Create the scatterplot
scatter_plot <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(x = "Time (Hour)", y = "Temperature", title = "Scatterplot of Temperature vs. Time (Hour)") +
  theme_minimal()

# Display the scatterplot
print(scatter_plot)
```

In the scatterplot we observe that during the month of January the temperature at the beginning of the month was around 40 degrees fahrenheit and sharply decreased to below 25 degrees. The temperature fluctuates up and down as the days cycle through day and night. While there are fluctuations, the temperature appears to follow a generally increasing trend throughout the month until January 15th where there is a peak temperature of around 60 degrees, followed by a sharp decline down to about 35 degrees where the temperature remained for the rest of the month. There was a notable period of high humidity, as shown by the lighter blue data points on the scatterplot, which started around January 10 and ended around the 15th. The humidity during the month of January generally hovered around 40-60%, and the humidity was high on the days where the temperature averages were higher and the temperature spread was smaller.

```{r}
# Export the scatterplot as an image file to project directory
ggsave("scatterplot_HW1_vsm2118.png")
```

# Problem 2 

```{r}
# Loading the tidyverse package
library(tidyverse)

# Create the data frame
my_data <- data.frame(
  numeric_var = rnorm(10),                   # Random sample from a standard Normal distribution
  logical_var = rnorm(10) > 0,              # Logical vector indicating whether elements are > 0
  character_var = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),  # Character vector
  factor_var = factor(rep(1:3, length.out = 10))  # Factor vector with 3 different levels
)
```

```{r}
# Attempt to calculate the mean for each variable
mean_numeric <- mean(pull(my_data, numeric_var))
  print(mean_numeric)
mean_logical <- mean(pull(my_data, logical_var))
  print(mean_logical)
mean_character <- mean(pull(my_data, character_var))
  print(mean_character)
mean_factor <- mean(pull(my_data, factor_var))
  print(mean_factor)
```

#### What works and what doesnt? 
When taking the mean of the numeric and logical variables a numerical result was returned, however the factor and character variables produced warning messages when trying to take the means of the dataframe. Athough, this is expected given that calculating means for character and factor variables doesn't make sense in the traditional statistical sense because these variable types represent categorical data rather than numeric. 

```{r}
# Applying as.numeric to the variables
numeric_logical <- as.numeric(my_data$logical_var)
numeric_character <- as.numeric(my_data$character_var)
numeric_factor <- as.numeric(my_data$factor_var)
```

#### What happens, and why? Does this help explain what happens when you try to take the mean?
For the Logical Variable as.numeric converted TRUE to 1 and FALSE to 0 as expected without any error messages. However, the character variable resulted in a Warning message since as.numeric cannot directly convert characters like "A" to numeric, so it returns NA. For the factor variable, as.numeric attempted to convert factor levels to numeric, but the resulting values may not represent the original data's meaning, resulting in the Warning message and leading to NA. This helps to explain why taking the mean of non-numeric variables results in NA. It is because R requires numeric values to complete a mean calculation, and thus NA is returned when the data contains non-numeric values introduced by coercion. 
